An Analysis of the Feasibility and Performance of BitTorrent as a Data Transfer Protocol in Peer to Peer Backup Systems

### 2014-04-15 paper/paper.tex
improved upon methodology

\begin{titlepage}
\begin{center}
\thispagestyle{empty}
% TODO maybe nice pic? \includegraphics [width = 0.3\textwidth] {images/wpilogo}\\[0.3cm]
% TODO install \textsc font
\Large{\textbf{A Major Qualifying Project Report\\ \large{ON}}}\\[0.7cm]
\LARGE{\textsc {\textbf{An Analysis of the Feasibility and Performance of BitTorrent as a Data Transfer Protocol in Peer to Peer Backup Systems}}}\\[0.5cm]
\vspace{0.2cm}
\Large{\textbf{\\Submitted to the Faculty of }}
\LARGE{\textbf{\\WORCESTER POLYTECHNIC INSTITUTE\\}}
\vspace{0.5cm}
\Large{\textbf{\\In Partial Fulfillment of the Requirement for the}}
\Large{\textbf{\\Degree of Bachelor of Science}}
\vspace{0.5cm}
\Large{\textbf{\\by}}\\[0.5cm]
\large{\textbf{Dan Bouffard}}\\
\large{\textbf{Marc Green}}\\
\vspace{0.5cm}
\large{\textbf{UNDER THE GUIDANCE OF}}\\
\large{\textbf{Professor G\'abor N. S\'ark\"ozy}}\\
\vspace{1cm}
\large{\textbf{\\May 1, 2014}}\\
\end{center}
\end{titlepage}

@@

However, there are also several problems with using a centralized system to back up data. One of these problems is that there is a single point of failure, should the storage site be compromised in some way (for example, natural disasters). This is less of a problem for big companies with the resources to handle these situations, but there is still a cost to take the necessary preventative measures when creating the system.

@@

% TODO: reliable sources for data caps and monthly/yearly fees
Finally, there are often restrictions on what can be stored using centralized backup systems. This includes both the size of individual files and the amount of total data. More space is gained by paying a monthly or yearly fee, but even then the amount of space is capped at a specific size. For users who need to back up a large amount of large files, these systems are not very useful.

@@

Peer to Peer (P2P) Backup Systems are applications that allow users to back up data without the use of a central entity. Instead, each user offers some of his or her storage space in order to use the space of others as backup locations. We start this section by introducing peer to peer networks, and then discuss the technical components of existing P2P backup systems.

@@

A peer to peer network is a type of distributed network model in which participants form direct connections to each other. This is in contrast to the centralized client-server network model, in which all participating clients connect to a central server to carry out their task. Client-server network models are used on the World Wide Web: the clients are users' web browsers and the server is the website being visited. In peer to peer networks, each participant, or "peer", functions as both a client and a server; peers initiating a request take on the role of the client, and peers answering the request take on the role of the server.

In the client-server model, servers are expected to always be available and accessible. Peer to peer networks do not have this luxury. Instead, they experience the constant fluctuation of peers joining and leaving the network. This is called \textit{churn} \cite{StorageSearchP2PNetworks}. Churn in P2P networks complicates the communication between peers because the sought peer may be offline. This is especially a problem in P2P backup systems; for example, a peer might try to recover data from another peer who is not online. The strategies for dealing with churn, and many other P2P-specific complications, heavily depend on the type of overlay network a P2P application uses. We discuss churn in more detail in section \ref{sec:churn}.

@@

An overlay network is any network built on top of another network. That is, an overlay network consists of nodes with neighbors who are not physically connected, but instead use the underlying network to establish a virtual connection. They need not be constrained by the limitations of the physical world. Consider a trivial example where there are three computers connected to the Internet, but located in different parts of the world. These three computers could form an overlay network by agreeing to be virtual neighbors to each other. If their overlay network was used in a P2P application, then, when connecting to each other, they will use the underlying network, the Internet, to actually send messages. The P2P application does not need to know about the underlying network that is used to route messages; it only needs to know that there is an overlay network that defines the direct connections between peers.

Peer to peer applications use overlay networks as their foundation.\footnote{Overlay networks are used for many purposes, not just P2P applications. Note, however, in discussing overlay networks, we will still refer to their participants as \textit{peers}.}  The P2P overlay network provides structure for the application that uses it, allowing peers to find each other.

@@

\subsubsection{Structured Overlay Networks} \label{subsubsec:StructuredOverlayNetworks}

@@

To join a structured overlay network, a peer must know at least one other peer already in the network. This peer is often called the \textit{bootstrap peer}. The joining peer will contact the bootstrap peer to find its proper place in the topology. We come back to this below in more detail in our explanation of how to find other peers.

% TODO cite router.bittorrent.com as well known bootstrap node?
%In BitTorrent's structured overlay network \footnote{To clarify, we are referring to BitTorrent's distributed hash table, not its unstructured network. The latter involves using centralized trackers, the former removes such dependency. We refraim from using the term \textit{distributed hash table} to keep the explanation simple without loss of generality. We discuss distributed hash tables in section \ref{sec:backgroundDHT}.}, a well known bootstrap node \footnote{The BitTorrent specification uses \textit{node} to refer to participants of its structured overlay network and \textit{peer} as a more general term \cite{bittorrentDHT}. We will adopt this convention when speaking about BitTorrent, and use them interchangeably otherwise.} is router.bittorrent.com. A BitTorrent node randomly chooses a 160 bit \textit{nodeID} to use as an address. To join the network, it will contact a bootstrap node to find the nodes it is closest to in the topology. Closeness, in BitTorrent's case, is determined by taking the $XOR$ of \textit{nodeID}s; smaller values are defined as closer \cite{bittorrentDHT}.
% TODO define XOR?
% TODO maybe remove this entire paragraph, since it doesn't really clarify anything. it just gives another example of picking an id and mysteriously finding nodes its closest to. if we end up remove it, we should reword end of previous paragraph
% TODO maybe keep it in b/c we refer to closness in 3.1.1

@@

To find other peers, a peer in a structured overlay network will consult its routing table. This table is initially populated upon joining the network, and is kept up to date as peers join and leave. When a joining peer first contacts a bootstrap peer, it is unlikely that the bootstrap peer has the joining peer's direct neighbors in its routing table. This is because the routing table is designed to include many peers that are close, and only a few peers that are far\footnote{This is to provide an efficient, $O(\log n)$ time complexity. This means that given n nodes, it will take less than $c \times \log n$ steps (for some $c$), which is good performance.} A joining peer, by asking the bootstrap node, will learn of the peers in the bootstrap peer's routing table that the joining node is close to. The joining peer will then contact these peers, ask for the same information, and learn of peers it is even closer to. This process repeats until the joining node learns its direct neighbors \cite{p2pSurvey}.

Overlay networks are the abstraction that connects all the users of a P2P system. It is the foundation on which a P2P application can be built, and is responsible for providing methods for peers to join and leave the network, and find other peers in the network. Furthermore, structured overlay networks allow the P2P application to locate specific information very quickly because their virtual address space provides full accountability of any piece of data. It is for this reason we chose a structured overlay network as the substrate of our P2P application; it is easy to determine which peer is responsible for a given backup because we know it will be the one closest in the address space.\footnote{However, churn complicates this. For example, a peer with an even closer address may join the network between storing and retrieving the data.}

% TODO may need better segue between p2p networks and file chunking

% TODO: Ask Sarkozy/Mihaly if they think keeping this section in is worth it.
\subsection{File Chunking} \label{sec:filechunking}
The format in which files are stored can have an effect on the efficiency of the P2P backup system. At first, backing up each file as a whole might seem like the obvious choice, since files are only useful in their entirety. However, the creators of the P2P backup systems pStore \cite{pStore} and PeerStore \cite{PeerStore} have shown that there are benefits to splitting files up into chunks, rather than distributing complete files to peers. In these systems, this is accomplished by creating two types of files: a "file block", which contains a section of the original file, and a "file block list", which contains information on which file blocks compose the original file and how to order them. This has the benefit of having data units be a fixed size that will be easy to store for any node. For example, if a peer needs to store a 10GB file, it does not need to find a peer capable of storing 10GB (which might be rare); rather, it needs to find $file\_size/chunk\_size \times number\_of\_replicas$ peers. Although lowering the chunk size will increase the chances of finding a peer to store on, it is important to choose a reasonable chunk size, as using more nodes for storage will add more overhead in finding those nodes.

@@

Our system does not use file chunking. Instead, by using BitTorrent as the data transport protocol, we have applied the concept of splitting files into chunks in a new way in P2P backup systems. We distribute complete files to each backup location, but separate files into chunks when they are being transferred; this is how BitTorrent works. With the BitTorrent Protocol, a peer can download a file by requesting chunks of it from many peers at the same time. By utilizing the upload bandwidth of several peers, a file can be downloaded quicker than if it were directly downloaded from one entity (see section \ref{sec:TheBitTorrentProtocol} for more details).

\subsection{Fairness} \label{sec:BackgroundFairness}
Another aspect of P2P backup systems is fairness between peers. Peers need to be storing an amount of data proportional to the amount that they are backing up to other peers. This can be challenging to enforce, because untrustworthy peers can attempt to game the system to obtain free backup storage. For example, consider a simple system where each peer must offer to store the same amount of data that they back up on other peers. A peer could claim that all of the storage space that it should be offering is being used by other peers, while in reality the peer has not taken in any other peer's data. It is clear that a mechanism needs to be built into the system that ensures that peers store the data of each other.

PeerStore handles this via a combined trading and challenge system. When a peer wants to create a backup of its data, it finds a peer to trade data with. In its request to back up data, the peer will also advertise that it is willing to store a certain amount of data for its trading partner. The peer that receives the request will then need to accept or reject it, and if the request is accepted, the acceptor will also need to request a certain amount of storage in return. If the original requestor accepts this offer, then the two become trading partners and will send each other data. This process is illustrated in figure \ref{fig:peerStoreTrading} \cite{PeerStore}.

\begin{figure}
  \centering
  \includegraphics[scale=0.75]{PeerStoreTrading}
  \caption{Trading protocol of PeerStore \cite{PeerStore}}
  \label{fig:peerStoreTrading}
\end{figure}

While this ensures that peers will agree to take data in proportion to the amount that they want to back up, it still does not guarantee that peers hold onto that data. For example, a peer could form a trading agreement with another peer, exchange data, and promptly discard the received data. To combat this, peers can periodically challenge each other to prove that they are actually storing data. This is the approach that we, like PeerStore, propose. We provide peers the ability to ask each other to prove they are storing the data that they are supposed to be storing. A peer proves itself by replying with a value that could only be derived if they were, indeed, safekeeping the data. See \ref{subsec:ChallengeMechanism_sec:Fairness_chap:BTBackup}
 for more details about our proposed challenge mechanism.

\subsection{Churn and Data Migration} \label{sec:churn}
With each storage node being regular users, it is often the case that nodes can be offline. Recall that churn refers to the fact that peers will continuously enter and leave the P2P network. The designer of a P2P backup system must take this into consideration if the system is to provide reliable access to backed up data. This has been accomplished via data migration \cite{pStore,PeerStore}, where the availability of files is periodically checked and replicas are made when the peers being used for backup are not reliably available. Other data migration algorithms have been designed specifically to work in high churn networks \cite{StorageSearchP2PNetworks}. In this case, data replicas are periodically redistributed to different groups of nodes, which means that newer nodes have a chance to store data. This is beneficial because data stays evenly distributed throughout the overlay network, preventing any single node from being responsible for too much data, regardless of when they joined.\footnote{It is important to note this strategy is designed to work in an unstructured overlay network, not a structured one. Structured networks need not redistribute like this because the evenly distributed address assignment provides the even data distribution.}

The number of replicas to maintain will have an impact on the performance of the system. If the replica count is too low, the probability of recovering the  file is lowered. If the replica count is too high, then storage space and bandwidth will be needlessly used. Furthermore, if the system requires a peer to offer a proportional amount of storage space to the amount that it is using on others, the peer will also have to offer a lot more space than what it is using locally for its own data. % TODO this last sentence might go better in the fairness section

\subsection{Cryptography} \label{sec:crypto}
Crytography is the study of techniques for secure communication in the presence of third parties \cite{cryptoDef}. It is heavily relied upon in the computer world to protect secrets, verify identities, and ensure data hasn't been tampered with. Our system uses cryptography to \textit{encrypt} users' data before replicating it to stranger's computers. Encryption is the process of obfuscating text into a state such that it cannot be understood without reversing the process. There are several types of encryption proccesses, each with its respective \textit{decryption} process. Decryption refers to reversing the encryption so that encrypted message becomes readable again.

@@

Encryption in P2P backup systems is necessary to prevent the peers who are storing backups of data from reading it. A system in which backups were unencrypted would be much less popular with users, because they would lose privacy by using it. The terminology associated with this is \textit{confidentiality}, meaning that data that is supposed to be confidential, remains confidential.

\textit{Hashing} is the process of applying a \textit{hash function}, which uses a mathematical function, to some arbitrary-length message, resulting in a fixed sized \textit{hash}. Regardless of the input, the hash will usually look like a random string of characters because the mathematical function manipulates the underlying data. This manipulation greatly changes the bits originally used to represent the input data. A good hash function will uniformly distribute the hashes it produces to minimize \textit{collisions}. Collisions are when two inputs hash to the same value, and are often undesired, especially in \textit{cryptographic hash functions}. Cryptographic hash functions are hash functions that are suitable for use in cryptography. They possess mathematical properties that allow for confidentiality to be provided, among other properties.

To strengthen hashing algorithms, a \textit{salt} can be utilized. A salt is essentially a randomly picked string of bytes that is concatenated to data to be hashed. The use of different salts with the same sequence of data creates a different hash (since the total input the hash function is different). The same hash will be calculated if the same salt is used with the same sequence of data. Using salts is a useful technique in preventing \textit{dictionary attacks}, where an attacker compares a precomputed list of inputs and their hashes with a target hash of some unknown data in an attempt to find out what the unknown data is. Dictionary attacks are common on password hashes, because, upon success, the attacker will discover the password used to generate the password hash. With a salt, the precomputed hash for the data wll not match. The list of hashes cannot be easily recomputed with the salt, since hashing is designed to be an expensive and lengthy operation.

@@

While this is a convenient optimization, it has a security weakness: any user who has access to the unencrypted version of a file can confirm whether or not that file is being stored in the system. This encroaches on users' privacy.

@@

A \textit{hash table} is a data structure that maps unique keys to values. It has attractive performance features, such as $O(1)$ average lookup time. This means that, on average, the time it takes to retrieve a value from the hash table is independent of the number of entries being stored. This performace comes from the use of a hash function when organizing the data internally, the details of which is out of the scope of this paper. Hash tables also have $O(1)$ insertion and deletion time.

A \textit{distributed hash table} (DHT) is a hash table that is stored across mutliple nodes. Each node is responsible for a portion of the hash table's contents. In exchange for distributing the storage cost among several nodes, DHTs do not have $O(1)$ search, insertion, or deletion time. However, this is generally acceptable because the usefulness of distributing the storage cost far outweighs the drawback of slower operations on the data structure. This is especially true since the network latency will likely overshadow the slower performance of the data structure. The alternative would be to have each peer keep a copy of the entire hash table. This solution suffers from several problems: it makes the hash table hard to maintain, the download time can become nontrivial, and it might become infeasible to store the entire table on regulars peers, since it can become huge with larger networks.

@@

DHTs play a major role in many P2P systems. They serve as the means through which peers can find either data or other peers, depending on the system; our system uses a DHT for the latter. In conjunction with the structured overlay network, the DHT allows for peers to discover each other. This is primarily done when looking for peers to send backups to. We discuss how we use a DHT in more detail in section \ref{sec:DataExchange}.

\section{The BitTorrent Protocol} \label{sec:TheBitTorrentProtocol}
BitTorrent is a peer to peer protocol designed for fast file sharing. A peer who wishes to share a file creates a .torrent file containing metadata about the file to be shared. This peer is known as the file's initial \textit{seeder}, or uploader, and through the use of a BitTorrent client, starts \textit{seeding}, or uploading, the file. When a peer acts as a seeder, its sole purpose is to share the file that it is seeding with other peers. Peers who wish to download the file must acquire its .torrent file through an out-of-band channel, which is often a torrent indexer. Peers use the .torrent file to locate the initial seeder through either a centralized tracker or a DHT. These peers then connect to the initial seeder and start downloading the file. These peers are known as \textit{leechers} until they possess a full copy of the file, which is when they, themselves, become seeders.

% TODO: add citation and more information about how much better BitTorrent is vs. direct download. Actually, add a lot of citations to this paragraph
While this is more complex than hosting a file on a server and having users download it, with enough peers there is a huge performance gain with respect to download speeds. The reason this works is due to the differences in upload and download speeds on the Internet; with most Internet connections offered to consumers, the upload bandwidth is much smaller than the download bandwidth. This means that, in the scenario where one machine is tranferring data over the Internet to another machine, download speeds would be severely limited by the upload bandwidth of the uploader. If multiple machines are trying to download from the same uploader, the download speeds will be even worse, as the upload bandwidth is being shared.

% TODO: finish this paragraph
The BitTorrent Protocol takes advantage of this by having leechers download content from multiple peers who are a part of the torrent. A file is partitioned into several \textit{chunks}, or file blocks, and a file download is performed by requesting these chunks from multiple peers. % Use combined upload bandwidth, etc.

BitTorrent can transfer files quickly due to multiple seeders each uploading different chunks of the file to each leecher. This parallelized download is enhanced by the fact that leechers, too, upload the chunks of the file they possess to other leechers who do not have those chunks. \footnote {This is a simplified version of the BitTorrent protocol. For more information, see \cite{bittorrentProtocol}}

\section{BitTorrent Sync}
BitTorrent Sync is a new application developed by BitTorrent, Inc. It is a P2P synchronization tool that allows users to share files between trusted devices. This is accomplished by generating a set of 20-byte "secrets" for each folder to share and distributing the secret to each trusted device. All data is encrypted using AES-128 in counter mode \cite{btsynctech}, the key of which is derived from the secret. There are three different secrets, each of which provide different permissions for the peer that is given the secret: the read/write secret, the read only secret, and the encryption secret, which allows the node to store an encrypted version of the data so that it cannot read the contents. A variation of the read/write and read only secrets also exists called the one-time secret, which allows a peer to send out a valid key that is only usable by one node \cite{btsyncuserguide}.

@@

BitTorrent Sync leverages the BitTorrent protocol to provide quick downloads between synced peers. We, too, leverage this through the use of BitTorrent Sync to provide fast recovery of backed up files. The specification of our system is described in the next chapter.

@@

In this chapter, we discuss the design of our Peer to Peer backup system. See Chapter \ref{chap:impl} for our implementation of it.

\section{Data Exchange} \label{sec:DataExchange}

% todo reword
This section discusses the infrastructure and the protocols of our system that are concerned with moving data between entities in the network. Henceforth, we use the following terminology to disambiguate participants:

% todo use package enumitem to make labels italics instead of bold
\begin{description}
  \item[Peer:] A particpant who is sending replicas of their data into the network for safekeeping. We refer to this process as \textit{backing up} data.
  \item[Node:] A participant who is storing a replica for a peer. We refer to this process as \textit{storing} data.
\end{description}

This terminology is used to give context. It is very likely that every participant in the system will eventually serve both roles (perhaps even simultaneously), but we refer to them as \textit{peer} or \textit{node} depending on what role they are playing in a given situation. A peer plays the role of the client, whereas a node plays the role of a server.\footnote{We may qualify each term to help give context. For example, we may say "\ldots a \textit{replication} node \ldots" to emphasize the fact that the node is storing a replica.} When there is no context, we use the terms interchangeably. 
% TODO find/replace all uses of replicant node with replication node or vice versa

>>>>

\begin{titlepage}
\begin{center}
\thispagestyle{empty}
% TODO maybe nice pic? \includegraphics [width = 0.3\textwidth] {images/wpilogo}\\[0.3cm]
% TODO install \textsc font
\Large{\textbf{A Major Qualifying Project Report\\ \large{ON}}}\\[0.7cm]
\LARGE{\textsc {\textbf{An Analysis of the Feasibility and Performance of BitTorrent as a Data Transfer Protocol in Peer to Peer Backup Systems}}}\\[0.5cm]
\vspace{0.2cm}
\Large{\textbf{\\Submitted to the Faculty of }}
\LARGE{\textbf{\\WORCESTER POLYTECHNIC INSTITUTE\\}}
\vspace{0.5cm}
\Large{\textbf{\\In Partial Fulfillment of the Requirement for the}}
\Large{\textbf{\\Degree of Bachelor of Science}}
\vspace{0.5cm}
\Large{\textbf{\\by}}\\[0.5cm]
\large{\textbf{Dan Bouffard}}\\
\large{\textbf{Mark Green}}\\
\vspace{0.5cm}
\large{\textbf{UNDER THE GUIDANCE OF}}\\
\large{\textbf{Professor G\'abor N. S\'ark\"ozy}}\\
\vspace{1cm}
\large{\textbf{\\May 1, 2014}}\\
\end{center}
\end{titlepage}

@@

However, there are also several problems with using a centralized system to backup data. One of these problems is that there is a single point of failure, should the storage site be compromised in some way (for example, natural disasters). This is less of a problem for big companies with the resources to handle these situations, but there is still a cost to take the necessary preventative measures when creating the system.

@@

% TODO: reliable sources for data caps and monthly/yearly fees
Finally, there are often restrictions on what can be stored using centralized backup systems. This includes both the size of individual files and the amount of total data. More space is gained by paying a monthly or yearly fee, and even then the amount of space is capped at a specific size. For users who need to backup a large amount of large files, these systems are not very useful.

@@

Peer to Peer (P2P) Backup Systems are networks that allow users to backup data without the use of a central entity. Instead, each user offers some of his or her storage space in order to use the space of others as backup locations. We start this section by introducing peer to peer networks, and then discuss the technical components of existing P2P backup systems.

@@

A peer to peer network is a distributed network model in which participants form direct connections to each other. This is in contrast to the centralized client-server network model, in which all participating clients connect to a central server to carry out their task. Client-server network models are used on the World Wide Web: the clients are users' web browsers and the server is the website being visited. In peer to peer networks, each participant, or "peer", functions as both a client and a server; peers initiating a request take on the role of the client, and peers answering the request take on the role of the server.

In the client-server model, servers are expected to be always available and accessible. Peer to peer networks do not have this luxury. Instead, they experience the constant fluctuation of peers joining and leaving the network. This is called \textit{churn} \cite{StorageSearchP2PNetworks}. Churn in P2P networks complicates the communication between peers because the sought peer may be offline. This is especially a problem in P2P backup systems, for example, when trying to recover data from a peer who is not online. The strategies for dealing with churn, and many other P2P-specific complications, heavily depend on the type of overlay network a P2P application uses. We discuss churn in more detail in section \ref{sec:churn}.

@@

An overlay network is a network built on top of another network. That is, an overlay network consists of nodes with neighbors who are not physically connected, but instead use the underlying network to establish a virtual connection. They need not be constrained by the limitations of the physical world. Consider a trivial example where there are three computers connected to the Internet, but located in different parts of the world. These three computers could form an overlay network by agreeing to be virtual neighbors to each other. If their overlay network was used in a P2P application, then, when connecting to each other, they will use the underlying network, the Internet, to actually send messages. The P2P application does not need to know about the underlying network that is used to route messages, it only needs to know that there is an overlay network that defines the direct connections between peers.

Peer to peer applications use overlay networks as their foundation.\footnote{Overlay networks are used for many purposes, not just P2P applications. Note, however, in discussing overlay networks, we will refer to their participants as \textit{peers}.}  The P2P overlay network provides peers the ability to join and leave the network and find other peers with whom to communicate.

@@

\subsubsection{Structured Overlay Networks}

@@

To join a structured overlay network, a peer must know at least one other peer already in the network. This peer is often called the \textit{bootstrap peer}. The joining peer will contact the bootstrap peer to find its proper place in the topology. We address this in more detail in the explanation of how to find other peers.

% TODO cite router.bittorrent.com as well known bootstrap node?
%In BitTorrent's structured overlay network \footnote{To clarify, we are referring to BitTorrent's distributed hash table, not its unstructured network. The latter involves using centralized trackers, the former removes such dependency. We refraim from using the term \textit{distributed hash table} to keep the explanation simple without loss of generality. We discuss distributed hash tables in section \ref{sec:backgroundDHT}.}, a well known bootstrap node \footnote{The BitTorrent specification uses \textit{node} to refer to participants of its structured overlay network and \textit{peer} as a more general term \cite{bittorrentDHT}. We will adopt this convention when speaking about BitTorrent, and use them interchangeably otherwise.} is router.bittorrent.com. A BitTorrent node randomly chooses a 160 bit \textit{nodeID} to use as an address. To join the network, it will contact a bootstrap node to find the nodes it is closest to in the topology. Closeness, in BitTorrent's case, is determined by taking the $XOR$ of \textit{nodeID}s; smaller values are defined as closer \cite{bittorrentDHT}.
% TODO define XOR?
% TODO maybe remove this entire paragraph, since it doesn't really clarify anything. it just gives another example of picking an id and mysteriously finding nodes its closest to. if we end up remove it, we should reword end of previous paragraph

@@

To find other peers, a peer in a structured overlay network will consult its routing table. This table is initially populated upon joining the network, and is kept up to date as peers join and leave. When a joining peer first contacts a bootstrap peer, it is unlikely that the bootstrap peer has the joining peer's direct neighbors in its routing table. This is because the routing table is designed to include many peers that are close, and only a few peers that are far\footnote{This is to provide an efficient, $O(\log n)$ time complexity. This means that given n nodes, it will only take $c \times \log n$ steps, which is good performance.} A joining peer, by asking the bootstrap node, will learn of the peers in the bootstrap peer's routing table that the joining node is close to. The joining peer will then contact these peers, ask for the same information, and learn of peers it is even closer to. This process repeats until the joining node learns its direct neighbors \cite{p2pSurvey}.

Overlay networks are the abstraction that connects all the users of a P2P system. It is the foundation on which a P2P application can be built, and is responsible for providing methods for peers to join and leave the network, and find other peers in the network. Furthermore, structured overlay networks allow the P2P application to locate specific information very quickly because their virtual address space provides full accountability of any piece of data. It is for this reason we chose a structured overlay network as the substrate of our P2P application; it is easy to determine which peer is responsible for a given backup because we know it will be the one closest in the address space.\footnote{However, churn complicates this because a peer with an even closer address may join the network between storing and retrieving the data.}

% TODO: Ask Sarkozy/Mihaly if they think keeping this section in is worth it.
\subsection{File Chunking} \label{sec:filechunking}
The format in which files are stored can have an effect on the efficiency of the system. At first, backing up each file as a whole might seem like the obvious choice, since files are only useful in their entirety. However, the creators of the P2P backup systems pStore and PeerStore have shown that there are benefits to splitting files up into chunks, rather than distributing complete files to peers. In these systems, this is accomplished by creating two types of files: a "file block", which contains a section of the original file, and a "file block list", which contains information on which file blocks compose the original file and how to order them. They also use special files, called File Block Lists in pStore and Descriptors in PeerStore, that list which blocks belong to the file and where to find them in the network. This has the benefit of having data units be a fixed size that will be easy to store for any node. For example, if a peer needs to store a 10GB file, it does not need to find a peer capable of storing 10GB (which might be rare); rather, it needs to find $file\_size/chunk\_size*number\_of\_replicas$ peers. Although lowering the chunk size will increase the chances of finding a peer to store on, it is important to choose a reasonable chunk size, as using more nodes for storage will add more overhead in finding those nodes.

@@

Our system does not use file chunking. Instead, we have applied the concept of splitting files into chunks in a new way in P2P backup systems. We distribute complete files to each backup location, but separate files into chunks when they are being transferred; this is how BitTorrent works. With the BitTorrent Protocol, a peer can download a file by requesting chunks of it from many peers at the same time. By utilizing the upload bandwidth of several peers, a file can be downloaded quicker than if it were directly downloaded from one entity (see section \ref{thebittorrentprotocol} for more details).

\subsection{Fairness}
Another aspect of P2P backup systems is fairness between peers. Peers need to be storing an amount of data proportional to the amount that they are backing up to other peers. This can be challenging to enforce, because untrustworthy peers can attempt to game the system to obtain free backup storage. For example, consider a simple system where each peer must offer store the same amount of data that they back up on other peers. A peer could claim that all of the storage space that it should be offering is being used by other peers, while in reality the peer has not taken in any other peer's data. It is clear that a mechanism needs to be built into the system that ensures that peers store the data of each other.

PeerStore handles this via a combined trading and challenge system. When a peer wants to create a backup of its data, it finds a peer to trade data with. In its request to backup data, the peer will also advertise that it is willing to store a certain amount of data for its trading partner. The peer that receives the request will then need to accept or reject it, and if the request is accepted, the acceptor will also need to request a certain amount of storage in return. If the original requestor accepts this offer, then the two become trading partners and will send each other data. This process is illustrated in figure \ref{fig:peerStoreTrading} \cite{PeerStore}.

\begin{figure} \label{fig:peerStoreTrading}
  \centering
  \includegraphics[scale=0.75]{PeerStoreTrading}
  \caption{Trading protocol of PeerStore}
\end{figure}

While this ensures that peers will agree to take data in proportion to the amount that they want to backup, it still does not guarantee that peers hold onto that data. For example, a peer could form a trading agreement with another peer, exchange data, and promptly discard the received data. To combat this, peers can periodically challenge peers to prove that they are actually storing data. This is the approach that we, like PeerStore, propose. We provide peers the ability to ask each other to prove they are storing the data that they are supposed to be storing. A peer proves itself by replying with a value that could only be derived if they were, indeed, safekeeping the data. See \ref{sec:challenge} for more details about our proposed challenge mechanism.

\subsection{Churn and Data Migration} \label{sec:churn}
With each storage node being regular users, it is often the case that nodes can be offline. Recall that churn refers to fact that peers will continuously enter and leave the P2P network. The designer of a P2P backup system must take this into consideration if the system is to provide reliable access to backed up data. This has been accomplished via data migration \cite{pStore,PeerStore}, where the availability of files is periodically checked and replicas are made when the nodes being used for backup are not reliably available. Other data migration algorithms have been designed specifically to work in high churn networks \cite{StorageSearchP2PNetworks}. In this case, data replicas are periodically redistributed to different groups of nodes, which means that newer nodes have a chance to store data. This is beneficial because data stays evenly distributed throughout the overlay network, preventing any single node from being responsible for too much data, regardless of when they joined\footnote{It is important to note this strategy is designed to work in an unstructured overlay network, not a structured one. Structured networks need not redistribute like this because the evenly distributed address assignment provides the even data distribution.}

The number of replicas to maintain will have an impact on the performance of the system. If the replica count is too low, the probability of recovery the entire file is lowered. If the replica count is too high, then storage space and bandwidth will be needlessly used. Furthermore, if the system requires a peer to offer a proportional amount of storage space to the amount that it is using on other nodes, the peer will also have to offer a lot more space than what it is using locally for its own data. % TODO this last sentence might go better in the fairness section

\subsection{Cryptography} \label{sec:crypto}
Crytography is the study of techniques for secure communication in the presence of third parties \cite{cryptoDef}. It is heavily relied upon in the computer world to protect secrets, verify identities, and ensure data hasn't been tampered with. Our system uses cryptography to \textit{encrypt} users' data before replicating it to stranger's computers. Encryption is the process obfuscating text into a state such that it cannot be understood without reversing the process. There are several types of encryption proccesses, each with its respective \textit{decryption} process. Decryption refers to reversing the encryption so that encrypted message becomes readable again.

@@

Encryption in P2P backup systems is necessary to prevent the peers who are storing backups of data from reading the data they are backing up. A system in which backups were unencrypted would be much less popular with users, because they would lose privacy by using it. The terminology associated with this is \textit{confidentiality}, meaning that data that is supposed to be confidential, remains confidential.

\textit{Hashing} is the process of applying a \textit{hash function} to some arbitrary-length message, resulting in a fixed sized \textit{hash}. A good hash function will uniformly distribute the hashes it produces. \textit{Cryptographic hash functions} are hash functions that are suitable for use in cryptography. They possess mathematical properties that allow for confidentiality to be provided, among other properties.

@@

While this is a convenient optimization, it has the following security weakness: any user who has access to the unencrypted version of a file can confirm whether or not that file is being stored in the system. This encroaches on users' privacy.

@@

A \textit{hash table} is a data structure that maps unique keys to values. It has attractive performance features, such as $O(1)$ average lookup time. This means that, on average, the time it takes to retrieve a value from the hash table is independent of the number of entries being stored. This performace comes from the use of a hash function when organizing the data internally. Hash tables also have $O(1)$ insertion and deletion time.

A \textit{distributed hash table} (DHT) is a hash table that is stored across mutliple nodes. Each node is responsible for a portion of the hash table's contents. In exchange for distributing the storage cost among several nodes, DHTs do not have $O(1)$ search, insertion, or deletion time. However, this is generally acceptable because the usefulness of distributing the storage cost far outweights the drawback of slower operations on the data structure. This is especially true since the network latency will likely overshadow the slower performance of the data structure.

@@

DHTs play a major role in many P2P systems. They serve as the means through which peers can find either data or other peers, depending on the system; our system uses a DHT for the latter. In conjunction with the structured overlay network, the DHT allows for peers to discover each other. This is primarily done when looking for peers to send backups to. We discuss how we use our DHT in more detail in section \ref{sec:ourDHT}.

\section{The BitTorrent Protocol} \label{sec:thebittorrentprotocol}
BitTorrent is a peer to peer protocol designed for fast file sharing. A peer who wishes to share a file creates a .torrent file containing metadata about the file to be shared. This peer is known as the file's initial \textit{seeder}, and through the use of a BitTorrent client, starts \textit{seeding} the file. When a peer acts as a seeder, its sole purpose is to share the file that it is seeding with other peers. Peers who wish to download the file must acquire its .torrent file through an out-of-band channel, often a torrent indexer. Peers use the .torrent file to locate the initial seeder through either a centralized tracker or a DHT. These peers then connect to the initial seeder and start downloading the file. These peers are known as \textit{leechers} until they possess a full copy of the file, which is when they, themselves, become seeders.

% TODO: add citation and more information about how much better BitTorrent is vs. direct download. Actually, add a lot of citations to this paragraph
While this is more complex than hosting a file on a server and having all users download it, with enough peers there is a huge performance gain with respect to download speeds. The reason this works is due to the differences in upload and download speeds on the Internet; with most Internet connections offered to consumers, the upload bandwidth is much smaller than the download bandwidth. This means that, in the scenario where one machine is tranferring data over the Internet to another machine, download speeds would be severely limited by the upload bandwidth of the uploader. If multiple machines are trying to download from the same uploader, the download speeds will be even worse, as the upload bandwidth is being shared.

% TODO: finish this paragraph
The BitTorrent Protocol takes advantage of this by having leechers download content from multiple peers who are a part of the torrent. A file a partitioned into several file blocks, and a file download is performed by requesting these blocks from multiple peers. % Use combined upload bandwidth, etc.

BitTorrent can transfer files quickly due to multiple seeders each uploading different \textit{chunks} of the file to each leecher. This parallelized download is enhanced by the fact that leechers, too, upload the chunks of the file they possess to other leechers who don't have those chunks. \footnote {This is a simplified version of the BitTorrent protocol. For more information, see \cite{bittorrentProtocol}}

\section{BitTorrent Sync}
BitTorrent Sync is a new application developed by BitTorrent, Inc. It is a P2P synchronization tool that allows users to share files between trusted devices. This is accomplished by generating a set of 20-byte "secrets" for each folder to share and distributing the secret to each trusted device. All data is encrypted using AES-128 in counter mode \cite{btsynctech}, the key of which is derived from the secret. There are three different secrets, each of which provide different permissions for the peer that is given the secret: the read/write secret, the read only secret, and the encryption secret, which allows the node to store an encrypted version of the data, so that it cannot read the contents. A variation of the read/write and read only secrets also exists called the one-time secret, which allows a peer to send out a valid key that is only usable by one node \cite{btsyncuserguide}.

@@

BitTorrent Sync leverages the BitTorrent protocol to provide quick downloads between synced peers. We also leverage this through the use of BitTorrent Sync to provide fast recovery of backed up files. The implementation of our system is described in the next section.

@@

In this chapter, we discuss the design of our Peer to Peer backup system. See chapter \ref{chap:impl} for our implementation of it.

\section{Data Exchange}

This chapter discusses everything in our system that is concerned with moving data between entities in the network.

PeerStore showed that separating the metadata layer from the data layer can result in significant bandwidth reduction. Our system separates the two layers as well: the metadata layer is responsible for ..

\subsection{The Metadata Layer}

% The DHT (what information would be stored in it)
